\documentclass[10pt,journal,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{amsfonts} % assumes amsfonts package installed
\usepackage{epstopdf}
\usepackage{color}
\usepackage{comment}
\usepackage{algorithm}
\usepackage{bbding}
\usepackage{algorithmic}
\usepackage[mathscr]{eucal}
%\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}  % Use Input in the format of Algorithm  
\renewcommand{\algorithmicensure}{\textbf{Output:}} % Use Output in the format of Algorithm  
\usepackage{mathrsfs}

%The self define macro
\include{def}
% The figures path
\graphicspath{{figures/}}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%=======================================================================================

\begin{document}

\title{Online Observability of \\ 
	Boolean Control Networks}
\author{Guisen~Wu,
	Liyun~Dai\Envelope,
	Zhiming~Liu,
	Taolue~Chen,
	Jun~Pang,
	Hongyang~Qu
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem
Guisen Wu, Liyun Dai and Zhiming Liu are with the School of Computer and Information Science, Southwest University
Chongqing, 400715 China. E-mail: \{wgs233,~dailiyun,~zhimingliu88\}@swu.edu.cn.
%
Taolue Chen is with the Department of Computer Science and Information Systems, Birkbeck, University of London. E-mail: taolue@dcs.bbk.ac.uk.
Jun Pang is with the Faculty of Science, Technology and Communication
and the Interdisciplinary Centre for Security, Reliability and Trust, University of Luxembourg. E-mail: jun.pang@uni.lu.
Hongyang Qu is with the Department of Automatic Control and Systems Engineering, University of Sheffield. E-mail: h.qu@sheffield.ac.uk.}
\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}
}
 
\maketitle

%=======================================================================================


\begin{abstract}
Identification problem is an important topic of Boolean control networks (BCNs). And a BCN is identifiable iff it satisfies two conditions, the first one is the controllability, and second one is the initial state of it can be determined even if the initial state can not be reset. There is a type of BCNs' observability was proposed to research the identification problem, it states that there is an input sequence that determines the initial state.
%\tl{here it is so abruptly to jump from identification to observability} 
%And the observability of BCNs is proposed to present whether the initial states of BCNs can be determined. 
But it does not capture the ability to determine the initial state of BCN whose initial state can be reset. In this paper we propose a new type of observability named online observability to capture this ability. Such that we can research the identification problem better. Moreover, we also propose determination algorithm for the online observability.
%Four different ways to determine initial states of Boolean control networks (BCNs) are proposed for different applications. Four types of \BCNs' observability present whether the initial states of \BCNs\ can be determined by these ways, respectively. Among the existing definitions of observability, the type-III observability is proposed to research the identification problem of \BCNs. In this paper, we propose a new observability named online observability to replace the type-III observability, such that we can research the the identification problem better. Moreover, we also propose determination algorithm for the online observability.
%\JP{I don't understand the last sentence.}

%For online observability, we can derive and decide input sequence one by one based on the feedback from the outputs of the network in every time step. %Moreover, we also propose determination algorithm for online observability.
%However, all of them are offline observability, meaning that the input sequence is given before studying the \BCN's initial state and the input sequence  remains unchanged during the process of determining. 
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
%\begin{IEEEkeywords}
%Boolean control networks, observability, algorithms
%\end{IEEEkeywords}


\IEEEpeerreviewmaketitle


\input{sec/introduction}
\input{sec/preliminaries}
\input{sec/onlineobservability}
\input{sec/determining}
%\input{sec/applications}
\input{sec/conclusions}
\bibliographystyle{IEEEtran} 
\bibliography{bcn} 
\input{sec/appendix}
\end{document}


